---
title: "Fifa Ranking"
author: "D.O"
date: "October 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align="center", fig.height=5.5, fig.width=6, collapse=TRUE, comment="", prompt=TRUE)
options(width=63)

library(dplyr)
library(ggplot2)
```


dplyr reference: https://dplyr.tidyverse.org/reference/index.html

#Read Data
Some relevant columns in the dataset: (copied fromhttps://www.kaggle.com/tadhgfitzgerald/fifa-international-soccer-mens-ranking-1993now/version/3)

Fédération Internationale de Football Association (FIFA)
International Men's Ranking (August 1993 - June 2018) Updated June 7th 2018

  *rank
  *country_full
  *country_abrv
  *total_points
  *previous_points
  *rank_change
  *cur_year_avg
  *cur_year_avg_weighted
  *last_year_avg
  *last_year_avg_weighted
  *two_year_ago_avg
  *two_year_ago_weighted
  *three_year_ago_avg
  *three_year_ago_weighted
  *confederation
  *rank_date
  
# Reading Data
```{r}
men_fifa_ranking <- read.csv("fifaranking.csv")
attach(men_fifa_ranking)
dim(men_fifa_ranking)
head(men_fifa_ranking)
str(men_fifa_ranking)
mfifa<- men_fifa_ranking[, c('rank','country_full','rank_change','previous_points', 'rank_date','confederation')]
head(mfifa)
tail(mfifa)
````

#Exploratory Data Analysis (EDA)
```{r}
nrow(mfifa)
ncol(mfifa)
country <- select(mfifa, country_full) %>% unique %>% nrow
table(mfifa$confederation) %>% barplot(col = "blue")

````

#Filter Confederation
```{r}
Confed <- mfifa$confederation
unique(Confed)
uefa <- filter(mfifa, confederation =='UEFA')
head(uefa,10)
afc <- filter(mfifa, confederation =='AFC')
head(afc,10)
caf <- filter(mfifa, confederation =='CAF')
head(caf,10)
conmebol <- filter(mfifa, confederation =='CONMEBOL')
head(conmebol,10)
concacaf<- filter(mfifa, confederation =='CONCACAF')
head(concacaf,10)
ofc<- filter(mfifa, confederation =='OFC')
head(ofc,10)
````
#Boxplots
```{r}
ggplot(data = mfifa, mapping = aes(x = confederation, y = rank, fill = factor(confederation))) + 
  geom_boxplot() +  coord_flip()
````


#Scatterplots
```{r}
ggplot(data = mfifa) +
  geom_point(mapping = aes(x = confederation, y = rank, color = confederation) )

````

#Top Ten from each Confederations
```{r}
top_uefa <- head(uefa,10)
top_afc <- head(afc,10)
top_caf <- head(caf,10)
top_conmebol <- head(conmebol,10)
top_concacaf <- head(concacaf,10)
top_ofc <- head(ofc,10)
top_conferation <- rbind(top_afc,top_caf,top_concacaf,top_conmebol,top_ofc,top_uefa)
top_conferation

ggplot(data = top_conferation, mapping = aes(x = confederation, y = rank, fill = factor(confederation))) + 
  geom_boxplot() +  coord_flip()

ggplot(data = top_conferation) +
  geom_point(mapping = aes(x = confederation, y = rank, color = confederation) )
````

